Interactively Visualizing Multivariate Market Segmentation Using pytourr

### 2024-09-25 manuscript.tex
caption for fig 4 and refs for intro

When there are more than two variables, histograms of the individual variables are commonly used to display the partitioning results. This means that the analyst likely cannot understand how the partitioning divides the data. All that they can observe is roughly how the individual variables relate to the partition, which is useful but inadequate. Here is where using tour methods to view high dimensions can be helpful. A tour (XXX general tour and recent tour) can be used to show scatterplots of combinations of variables and thus provide views like that in Figure \ref{kmeans-partition} where distinct differences between partitions can be observed. High dimensions are still tricky, and a combination of animations of the linear combinations, and interactive control (XXX REF radial/manual tour) over the combinations is important. A scatterplot of a combination of variables can be considered to be a projection of the data, and thus like a shadow of a 3D object, some aspects of the data (object) can be obscured. Using slices of the projected data (XXX REF) can be a useful addition to projections. This paper illustrates how to do this to better understand partitioning results for multivariate data.

@@

This paper is organised as follows XXX The methods are illustrated using Austrian and Australian tourism data provided in \cite{leisch2018market}.

\section{Interactive interface for partitioning}

@@

While tour animations are best obtained within R using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}. Interactive graphics are available when using Javascript, as implemented in the detourr package \citep{detourr}. This allows to replay a recorded tour path with interactive graphics, and can also be linked with additional displays, but lacks capabilities for manual tours.

@@

The interface was implemented in \texttt{pytourr} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives. The ability to navigate through various projections of the displayed tours directly within the GUI enables users to explore different aspects of the dataset. Furthermore, users can initiate new tours directly from the interface. The GUI also supports interactive variable selection, allowing users to specify which subset of variables should be visualized in the plots. Once users have identified interesting views or settings, \texttt{pytourr} allows them to save the displayed projections, subsets, and plots. This functionality ensures that analysis states can be preserved for further examination or reporting, making the package particularly useful for iterative analysis where findings may need to be revisited or shared with collaborators.

@@

\subsection{The graphical user interface}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{GUI_overview.png}
    \caption{Overview of the GUI, consisting of the sidebar (left) and display area (right). A: Sidebar. B: Display area. C: Feature selection checkboxes. D: Subset selection with color indicators. E: Frame selection interface. F: Save and load buttons. G: Interface for starting new tours. H: Metric selection interface.}
    \label{fig:GUI_overview}
\end{figure}

@@

\section{Workflow with the pytourr package}

@@

\section{Applications}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winteractiv_heatmap.png}
    \caption{A: heatmap displaying the intra cluster fraction. B: heatmap displaying the intra feature fraction.}
    \label{fig:winteractiv_heatmap}
\end{figure}

In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster. For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not. Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.

@@

\section{Discussion}

@@

\section{Conclusion}

@@

Future developments in the software might include expanding the range of interactive features and providing additional visualization methods.  Because it is built on R and python, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI. For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours, would be recommended.

>>>>

When there are more than two variables, histograms of the individual variables are commonly used to display the partitioning results. This means that the analyst likely cannot understand how the partitioning divides the data. All that they can observe is roughly how the individual variables relate to the partition, which is useful but inadequate. Here is where using tour methods to view high dimensions can be helpful. A tour (\cite{Asimov1985-xr}, and see \cite{lee2022}) can be used to show scatterplots of combinations of variables and thus provide views like that in Figure \ref{kmeans-partition} where distinct differences between partitions can be observed. High dimensions are still tricky, and a combination of animations of the linear combinations, and interactive control \citep{cook_manual_1997, laa2023new} over the combinations is important. A scatterplot of a combination of variables can be considered to be a projection of the data, and thus like a shadow of a 3D object, some aspects of the data (object) can be obscured. Using slices of the projected data \citep{Laa2020} can be a useful addition to projections. This paper illustrates how to do this to better understand partitioning results for multivariate data.

@@

This paper is organised as follows. Section \ref{interface} describes the software interface. The user workflow is explained in \ref{workflow}.  The methods are illustrated in Section \ref{applications} using Austrian and Australian tourism data provided in \cite{leisch2018market}. Sections \ref{discussion} and \ref{conclusion} discuss the limitations and potential future developments.

\section{Interactive interface for partitioning}~\label{interface}

@@

While tour animations are best obtained within R using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}. Interactive graphics are available when using Javascript, as implemented in the detourr package \citep{RJ-2023-052}. This allows to replay a recorded tour path with interactive graphics, and can also be linked with additional displays, but lacks capabilities for manual tours.

@@

The interface was implemented in \texttt{pytourr} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives. The ability to navigate through various projections of the displayed tours directly within a graphical user interface (GUI) enables users to explore different aspects of the dataset. Furthermore, users can initiate new tours directly from the interface. The GUI also supports interactive variable selection, allowing users to specify which subset of variables should be visualized in the plots. Once users have identified interesting views or settings, \texttt{pytourr} allows them to save the displayed projections, subsets, and plots. This functionality ensures that analysis states can be preserved for further examination or reporting, making the package particularly useful for iterative analysis where findings may need to be revisited or shared with collaborators.

@@

\subsection{Overview of the graphical user interface (GUI)}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{GUI_overview.png}
    \caption{Overview of the GUI. A: Sidebar controls; B: Display area;  C: Feature selection checkboxes; D: Subset selection with color indicators; E: Frame selection interface; F: Save and load buttons; G: Interface for starting new tours; H: Metric selection interface.}
    \label{fig:GUI_overview}
\end{figure}

@@

\section{Workflow with the pytourr package}~\label{workflow}

@@

\section{Applications}~\label{applications}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{winteractiv_heatmap.png}
    \caption{Traditional overview of clusters. Color represents  (A) the intra-cluster fraction, and (B) the intra-feature fraction, with lighter indicating higher values. From A, we can see that some activities dominate most clusters, e.g. relaxing and going for walks. Some activities are unique to clusters, e.g. skiing is a primary activity for clusters 1, 3, 5, 6. From B, we can see which activities dominate each cluster, e.g. cluster 4 tourists like going to a spa and using health activities.}
    \label{fig:winteractiv_heatmap}
\end{figure}

\textcolor{red}{It might be better to standardise these intra fractions. For A, one should divide by cluster count, and then compute activity fraction. For B, one should divide by activity count and then compute cluster fraction.} In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster. For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not. Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.

@@

\section{Discussion}~\label{discussion}

@@

\section{Conclusion}~\label{conclusion}

@@

Future developments in the software might include expanding the range of interactive features and providing additional visualization methods.  Because it is built on R and python, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI. For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended.
